# Reproducible Research: Peer Assessment 1
```{r echo=TRUE}
library('dplyr')
library('ggplot2')
```

## Loading and preprocessing the data
```{r echo=TRUE}
unzip('activity.zip', overwrite=TRUE)
df <- read.csv('activity.csv', stringsAsFactors=FALSE)
#df$date <- strptime(df$date, '%Y-%m-%d')
```

## What is mean total number of steps taken per day?
```{r echo=TRUE}
by.date <- group_by(df, date)
total.by.date <- summarise(by.date, steps=sum(steps))
qplot(x=total.by.date$steps, geom="histogram",
      xlab="Total Number of Steps Taken Each Day",
      ylab="Frequency",
      binwidth=1000)
mean.steps.per.day = mean(total.by.date$steps, na.rm=TRUE)
median.steps.per.day = median(total.by.date$steps, na.rm=TRUE)
```

The data has a `r mean.steps.per.day` mean total number of steps each day while
having a `r median.steps.per.day` median total number of steps each day.


## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
