# Reproducible Research: Peer Assessment 1
```{r echo=TRUE}
library('dplyr')
library('ggplot2')
```

## Loading and preprocessing the data
```{r echo=TRUE}
unzip('activity.zip', overwrite=TRUE)
df <- read.csv('activity.csv', stringsAsFactors=FALSE)
#df$date <- strptime(df$date, '%Y-%m-%d')
```

## What is mean total number of steps taken per day?
```{r echo=TRUE}
by.date <- group_by(df, date)
total.by.date <- summarize(by.date, steps=sum(steps, na.rm=TRUE))
qplot(x=total.by.date$steps, geom="histogram",
      xlab="Total Number of Steps Taken Each Day",
      ylab="Frequency",
      binwidth=1000)
mean.steps.per.day = mean(total.by.date$steps, na.rm=TRUE)
median.steps.per.day = median(total.by.date$steps, na.rm=TRUE)
```

The data has a `r mean.steps.per.day` mean total number of steps each day while
having a `r median.steps.per.day` median total number of steps each day.

## What is the average daily activity pattern?
```{r echo=TRUE}
by.interval <- group_by(df, interval)
average.by.interval <- summarize(by.interval, steps=mean(steps, na.rm=TRUE))
qplot(x=interval, y=steps, data=average.by.interval,
      geom="line",
      xlab="Interval",
      ylab="Mean Steps Across All Days")
max.average <- max(average.by.interval$steps)
max.average.interval = average.by.interval[max.average,
                                           'interval']
```
The `r max.average.interval` 5-minute interval has the maximum average steps
at `r max.average`.

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
